---
title: "Forecasting: Predicted Sales Data"
author: "Sonia Nikiema"
date: "12/22/2020"
output: 
  html_document: 
    toc: yes
---


# Reflection

Daily historical data is used to predict futur sales. This is kaggle competion and data comes from ["Predict Future Sales"](https://www.kaggle.com/c/competitive-data-science-predict-future-sales)

Let's get started!!!

# Forecasting 1C company total sales overtime

## Load Library

```{r Load library, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)  # For ggplot, dplyr, and friends
library(lubridate)
library(scales)     # For nicer labels
library(tsibble)
library(forecast)
library(tseries)  ##We will use tsible to create year_month index
library(zoo)
library(rio)
library(fable)
library(feasts)
library(stats)
```

## Load Package

```{r Load package}
# SalesTrain contains all necessary items for the prediction except ID #
raw_salesTrain<- read.csv("data/sales_train.csv")
```

## Big clean
### Clean sales_train


```{r}
salesTrain <- raw_salesTrain %>%
# converte date variable to date format
  mutate(date = gsub("[.]", "/", date))%>%
# converte date variable to date format
  mutate(date=as.Date.character(date, format="%d/%m/%Y"))%>%
  
# convert number variable into factor variable  
    mutate(shop_id=factor(shop_id))%>%
   mutate(item_id=factor(item_id))%>%
  mutate(month=factor(month(date),ordered = TRUE))%>%
  mutate(year=factor(year(date)))%>%

  # mutate item count day into item count month
   mutate(item_cnt_month=item_cnt_day)
```

### Select the predicted year

```{r}
clean_salesTrain<-salesTrain%>%
  filter(year==2015)%>%
  select(date, item_id, item_cnt_month)%>%
   mutate(year_month = yearmonth(date))
```

```{r Preapare data for time series plot}
  ggplot(clean_salesTrain, aes(x = year_month, y = item_cnt_month)) +
  geom_line()
```

Our data is now clean but it does mot have time-enabled data frame (a tsibble) which is indexed by the year+month for each row.

## Create time-enabled data frame

```{r}
library(tsibble)  # For embedding time things into data frames

# Since tsibble must have distinct rows identified by key and index. We will delete duplicates rows by grouping by year_month
sales_tsbl<-clean_salesTrain%>%
group_by(year_month)%>%
  summarise(item_cnt_month = n())%>%
  as_tsibble(index = year_month)
  
print(sales_tsbl)
```

## ARIMA() modeling in R
### Creating the model

```{r}
model <- sales_tsbl %>% 
  model(stl = STL(item_cnt_month ~ season(window="periodic")))

#check the model
model
```

### decompose the model

```{r}
decomp_model <- components(model)

#check decomposed model
decomp_model
```

### plot the model

```{r}
graph<- decomp_model %>%
  autoplot(season_adjust)
graph
```

```{r}
ggsave("graph.png", graph)
```


###  White Noise assumption 

```{r}
decomp_model %>%
  gg_tsdisplay(difference(season_adjust), plot_type='partial')
```

Obviously our data seems to be white noise and stationary. Therefore variables are independent and identically distributed with a mean of zero. This means that all variables have the same variance (sigma^2) and each value has a zero correlation with all other values in the series

###  fit the model

```{r}
# to avoid "Error: Failed to create output due to bad names. * Choose another strategy with `names_repair`", we will select the model components without .model

decomp_modelNostl<- decomp_model%>%
  select(-.model)

fit <- decomp_modelNostl %>%
  model(
    arima = ARIMA(season_adjust ~ pdq(0,0,0) + PDQ(0,0,0))
  )
report(fit)
```

###  checking residuals

```{r}
fit %>% 
  gg_tsresiduals()
```


### Forecast

```{r}
fit %>% 
  forecast(h=2) %>% 
  autoplot(decomp_modelNostl)
```


```{r}
fit %>% 
  forecast(h=2)
```

### Extract values from forecast to time series data frame

```{r}
tscountObject <- ts(sales_tsbl$item_cnt_month, 
  start = c(2015,1), 
  end = c(2015,10), 
  frequency = 10)

fcast<-forecast(arima(ts(tscountObject)))

sales_tsbl$forecast<- fcast$mean
sales_tsbl
```


```{r}
sim <- fit %>% 
  generate(h = 2, times = 5, bootstrap = TRUE)
sim
```







